% File borderland-15.txt
% Version 2017/09/18
% In the original, this was Chapter XXI.

\clearpage
\label{ch:15}

\begin{ChapterStart}
\null\null
\ChapterTitle{15. The Dark Sun}
\end{ChapterStart}

I was waked from my happiness, by a diminution of the pale and gentle light that lit the Sea of Sleep. I turned toward the huge, white orb, with a premonition of trouble. One side of it was curving inward, as though a convex, black shadow were sweeping across it. It was thus, that the darkness had come, before our last parting. I turned toward my Love, inquiringly. I noticed how wan and unreal she had grown, even in that brief space. Her voice seemed to come to me from a distance. The touch of her hands was no more than the gentle pressure of a summer wind, and grew less perceptible.

Already, quite half of the immense globe was shrouded. A feeling of desperation seized me. Was she about to leave me? Would she have to go, as she had gone before? I questioned her, anxiously, frightenedly; and she, nestling closer, explained, in that strange, faraway voice, that it was imperative she should leave me, before the Sun of Darkness---as she termed it---blotted out the light. At this confirmation of my fears, I was overcome with despair; and could only look, voicelessly, across the quiet plains of the silent sea.

At last, only a crescent of pale fire, lit the, now dim, Sea of Sleep. All this while, she had held me; but, with so soft a caress, that I had been scarcely conscious of it. We waited there, together, she and I; speechless, for very sorrow. In the dimming light, her face showed, shadowy---blending into the dusky mistiness that encircled us.

Then, when a thin, curved line of soft light was all that lit the sea, she released me---pushing me from her, tenderly. Her voice sounded, ‘I may not stay longer, Dear One.’ It ended in a sob.

She seemed to float away from me, and became invisible. Her voice came to me, out of the shadows, faintly; apparently from a great distance:---

‘A little while---’ It died away, remotely. In a breath, the Sea of Sleep darkened into night. Far to my left, I seemed to see, for a brief instant, a soft glow. It vanished, and, in the same moment, I became aware that I was no longer above the still sea; but once more suspended in infinite space, with the Green Sun---now eclipsed by a vast, dark sphere---before me.

Utterly bewildered, I stared, almost unseeingly, at the ring of green flames, leaping above the dark edge. Even in the chaos of my thoughts, I wondered, dully, at their extraordinary shapes. A multitude of questions assailed me. I thought more of her, I had so lately seen, than of the sight before me. My grief, and thoughts of the future, filled me. Was I doomed to be separated from her, always? Even in the old earth-days, she had been mine, only for a little while; then she had left me, as I thought, forever. Since then, I had seen her but these times, upon the Sea of Sleep.

A feeling of fierce resentment filled me, and miserable questionings. Why could I not have gone with my Love? What reason to keep us apart? Why had I to wait alone, while she slumbered through the years, on the still bosom of the Sea of Sleep? The Sea of Sleep! My thoughts turned, inconsequently, out of their channel of bitterness, to fresh, desperate questionings. Where was it? Where was it? I seemed to have but just parted from my Love, upon its quiet surface, and it had gone, utterly. It could not be far away! And the White Orb which I had seen hidden in the shadow of the Sun of Darkness! My sight dwelt upon the Green Sun---eclipsed. What had eclipsed it? Was there a vast, dead star circling it? Was the \textit{Central} Sun---as I had come to regard it---a double star? The thought had come, almost unbidden; yet why should it not be so?

My thoughts went back to the White Orb. Strange, that it should have been---I stopped. An idea had come, suddenly. The White Orb and the Green Sun! Were they one and the same? My imagination wandered backward, and I remembered the luminous globe to which I had been so unaccountably attracted. It was curious that I should have forgotten it, even momentarily. Where were the others? I reverted again to the globe I had entered. I thought, for a time, and matters became clearer. I conceived that, by entering that impalpable globule, I had passed, at once, into some further, and, until then, invisible dimension; There, the Green Sun was still visible; but as a stupendous sphere of pale, white light---almost as though its ghost showed, and not its material part.

A long time, I mused on the subject. I remembered how, on entering the sphere, I had, immediately, lost all sight of the others. For a still further period, I continued to revolve the different details in my mind.

In a while, my thoughts turned to other things. I came more into the present, and began to look about me, seeingly. For the first time, I perceived that innumerable rays, of a subtle, violet hue, pierced the strange semi-darkness, in all directions. They radiated from the fiery rim of the Green Sun. They seemed to grow upon my vision, so that, in a little, I saw that they were countless. The night was filled with them---spreading outward from the Green Sun, fan-wise. I concluded that I was enabled to see them, by reason of the Sun’s glory being cut off by the eclipse. They reached right out into space, and vanished.

Gradually, as I looked, I became aware that fine points of intensely brilliant light, traversed the rays. Many of them seemed to travel from the Green Sun, into distance. Others came out of the void, toward the Sun; but one and all, each kept strictly to the ray in which it traveled. Their speed was inconceivably great; and it was only when they neared the Green Sun, or as they left it, that I could see them as separate specks of light. Further from the sun, they became thin lines of vivid fire within the violet.

The discovery of these rays, and the moving sparks, interested me, extraordinarily. To where did they lead, in such countless profusion? I thought of the worlds in space.... And those sparks! Messengers! Possibly, the idea was fantastic; but I was not conscious of its being so. Messengers! Messengers from the Central Sun!

An idea evolved itself, slowly. Was the Green Sun the abode of some vast Intelligence? The thought was bewildering. Visions of the Unnameable rose, vaguely. Had I, indeed, come upon the dwelling-place of the Eternal? For a time, I repelled the thought, dumbly. It was too stupendous. Yet....

Huge, vague thoughts had birth within me. I felt, suddenly, terribly naked. And an awful Nearness, shook me.

And Heaven ...! Was that an illusion?

My thoughts came and went, erratically. The Sea of Sleep---and she! Heaven.... I came back, with a bound, to the present. Somewhere, out of the void behind me, there rushed an immense, dark body---huge and silent. It was a dead star, hurling onward to the burying place of the stars. It drove between me and the Central Suns---blotting them out from my vision, and plunging me into an impenetrable night.

An age, and I saw again the violet rays. A great while later---aeons it must have been---a circular glow grew in the sky, ahead, and I saw the edge of the receding star, show darkly against it. Thus, I knew that it was nearing the Central Suns. Presently, I saw the bright ring of the Green Sun, show plainly against the night The star had passed into the shadow of the Dead Sun. After that, I just waited. The strange years went slowly, and ever, I watched, intently.

The thing I had expected, came at last---suddenly, awfully. A vast flare of dazzling light. A streaming burst of white flame across the dark void. For an indefinite while, it soared outward---a gigantic mushroom of fire. It ceased to grow. Then, as time went by, it began to sink backward, slowly. I saw, now, that it came from a huge, glowing spot near the center of the Dark Sun. Mighty flames, still soared outward from this. Yet, spite of its size, the grave of the star was no more than the shining of Jupiter upon the face of an ocean, when compared with the inconceivable mass of the Dead Sun.

I may remark here, once more, that no words will ever convey to the imagination, the enormous bulk of the two Central Suns.

\clearpage
