% File borderland-18.txt
% Version 2017/09/18
% In the original, this was Chapter XXIV.

\clearpage
\label{ch:18}

\begin{ChapterStart}
\null\null
\ChapterTitle{18. The Footsteps in the Garden}
\end{ChapterStart}

Pepper is dead! Even now, at times, I seem scarcely able to realize that this is so. It is many weeks, since I came back from that strange and terrible journey through space and time. Sometimes, in my sleep, I dream about it, and go through, in imagination, the whole of that fearsome happening. When I wake, my thoughts dwell upon it. That Sun---those Suns, were they indeed the great Central Suns, ’round which the whole universe, of the unknown heavens, revolves? Who shall say? And the bright globules, floating forever in the light of the Green Sun! And the Sea of Sleep on which they float! How unbelievable it all is. If it were not for Pepper, I should, even after the many extraordinary things that I have witnessed, be inclined to imagine that it was but a gigantic dream. Then, there is that dreadful, dark nebula (with its multitudes of red spheres) moving always within the shadow of the Dark Sun, sweeping along on its stupendous orbit, wrapped eternally in gloom. And the faces that peered out at me! God, do they, and does such a thing really exist? ... There is still that little heap of grey ash, on my study floor. I will not have it touched.

Now that I am writing, let me record that I am certain, there is something horrible about to happen. Last night, a thing occurred, which has filled me with an even greater terror, than did the Pit. I will write it down now, and, if anything more happens, endeavor to make a note of it, at once. I have a feeling, that there is more in this last affair, than in all those others. I am shaky and nervous, even now, as I write. Somehow, I think death is not very far away.\linebreak Not that I fear death---as death is understood. Yet, there is that in the air, which bids me fear---an intangible, cold horror. I felt it last night. It was thus:---

Last night, I was sitting here in my study, writing. The door, leading into the garden, was half open. At times, the metallic rattle of a dog’s chain, sounded faintly. It belongs to the dog I have bought, since Pepper’s death. I will not have him in the house---not after Pepper. Still, I have felt it better to have a dog about the place. They are wonderful creatures.

I was much engrossed in my work, and the time passed, quickly. Suddenly, I heard a soft noise on the path, outside in the garden---pad, pad, pad, it went, with a stealthy, curious sound. I sat upright, with a quick movement, and looked out through the opened door. Again the noise came---pad, pad, pad. It appeared to be approaching. With a slight feeling of nervousness, I stared into the gardens; but the night hid everything.

Then the dog gave a long howl, and I started. For a minute, perhaps, I peered, intently; but could hear nothing. After a little, I picked up the pen, which I had laid down, and recommenced my work. The nervous feeling had gone; for I imagined that the sound I had heard, was nothing more than the dog walking ’round his kennel, at the length of his chain.

A quarter of an hour may have passed; then, all at once, the dog howled again, and with such a plaintively sorrowful note, that I jumped to my feet, dropping my pen, and inking the page on which I was at work.

‘Curse that dog!’ I muttered, noting what I had done. Then, even as I said the words, there sounded again that queer---pad, pad, pad. It was horribly close---almost by the door, I thought. I knew, now, that it could not be the dog; his chain would not allow him to come so near.

The dog’s growl came again, and I noted, subconsciously, the taint of fear in it.

Frightened, and puzzled, I seized a stick from the corner, and went toward the door, silently; taking one of the candles with me. I had come to within a few paces of it, when, suddenly, a peculiar sense of fear thrilled through me---a fear, palpitant and real; whence, I knew not, nor why. So great was the feeling of terror, that I wasted no time; but retreated straight-way---walking backward, and keeping my gaze, fearfully, on the door. I would have given much, to rush at it, fling it to, and shoot the bolts; for I have had it repaired and strengthened, so that, now, it is far stronger than ever it has been. I continued my, almost unconscious, progress backward, until the wall brought me up. At that, I started, nervously, and glanced ’round, apprehensively. As I did so, my eyes dwelt, momentarily, on the rack of firearms, and I took a step toward them; but stopped, with a curious feeling that they would be needless. Outside, in the gardens, the dog moaned, strangely.

Pad, pad, pad---Something passed down the garden path, and a faint, mouldy odor seemed to come in through the open door, and mingle with the burnt smell.

The dog had been silent for a few moments. Now, I heard him yowl, sharply, as though in pain. Then, he was quiet, save for an occasional, subdued whimper of fear.

A minute went by; then the gate on the West side of the gardens, slammed, distantly. After that, nothing; not even the dog’s whine.

I must have stood there some minutes. Then a fragment of courage stole into my heart, and I made a frightened rush at the door, dashed it to, and bolted it. After that, for a full half-hour, I sat, helpless---staring before me, rigidly.

Slowly, my life came back into me, and I made my way, shakily, up-stairs to bed.

\clearpage
