% File borderland-02.txt
% Version 2017/09/18
% In the original, this was Chapter III.

\clearpage
\label{ch:02}

\begin{ChapterStart}
\null\null
\ChapterTitle{2. The House in the Arena}
\end{ChapterStart}

And so, after a time, I came to the mountains. Then, the course of my journey was altered, and I began to move along their bases, until, all at once, I saw that I had come opposite to a vast rift, opening into the mountains. Through this, I was borne, moving at no great speed. On either side of me, huge, scarped walls of rocklike substance rose sheer. Far overhead, I discerned a thin ribbon of red, where the mouth of the chasm opened, among inaccessible peaks. Within, was gloom, deep and somber, and chilly silence. For a while, I went onward steadily, and then, at last, I saw, ahead, a deep, red glow, that told me I was near upon the further opening of the gorge.

A minute came and went, and I was at the exit of the chasm, staring out upon an enormous amphitheatre of mountains. Yet, of the mountains, and the terrible grandeur of the place, I recked nothing; for I was confounded with amazement to behold, at a distance of several miles and occupying the center of the arena, a stupendous structure built apparently of green jade. Yet, in itself, it was not the discovery of the building that had so astonished me; but the fact, which became every moment more apparent, that in no particular, save in color and its enormous size, did the lonely structure vary from this house in which I live.

For a while, I continued to stare, fixedly. Even then, I could scarcely believe that I saw aright. In my mind, a question formed, reiterating incessantly: ‘What does it mean?’ ‘What does it mean?’ and I was unable to make answer, even out of the depths of my imagination. I seemed capable only of wonder and fear. For a time longer, I gazed, noting continually some fresh point of resemblance that attracted me. At last, wearied and sorely puzzled, I turned from it, to view the rest of the strange place on to which I had intruded.

Hitherto, I had been so engrossed in my scrutiny of the House, that I had given only a cursory glance ’round. Now, as I looked, I began to realize upon what sort of a place I had come. The arena, for so I have termed it, appeared a perfect circle of about ten to twelve miles in diameter, the House, as I have mentioned before, standing in the center. The surface of the place, like to that of the Plain, had a peculiar, misty appearance, that was yet not mist.

From a rapid survey, my glance passed quickly upward along the slopes of the circling mountains. How silent they were. I think that this same abominable stillness was more trying to me than anything that I had so far seen or imagined. I was looking up, now, at the great crags, towering so loftily. Up there, the impalpable redness gave a blurred appearance to everything.

And then, as I peered, curiously, a new terror came to me; for away up among the dim peaks to my right, I had descried a vast shape of blackness, giantlike. It grew upon my sight. It had an enormous equine head, with gigantic ears, and seemed to peer steadfastly down into the arena. There was that about the pose that gave me the impression of an eternal watchfulness---of having warded that dismal place, through unknown eternities. Slowly, the monster became plainer to me; and then, suddenly, my gaze sprang from it to something further off and higher among the crags. For a long minute, I gazed, fearfully. I was strangely conscious of something not altogether unfamiliar---as though something stirred in the back of my mind. The thing was black, and had four grotesque arms. The features showed indistinctly, ’round the neck, I made out several light-colored objects. Slowly, the details came to me, and I realized, coldly, that they were skulls. Further down the body was another circling belt, showing less dark against the black trunk. Then, even as I puzzled to know what the thing was, a memory slid into my mind, and straightway, I knew that I was looking at a monstrous representation of Kali, the Hindu goddess of death.

Other remembrances of my old student days drifted into my thoughts. My glance fell back upon the huge beast-headed Thing. Simultaneously, I recognized it for the ancient Egyptian god Set, or Seth, the Destroyer of Souls. With the knowledge, there came a great sweep of questioning---‘Two of the---!’ I stopped, and endeavored to think. Things beyond my imagination peered into my frightened mind. I saw, obscurely. ‘The old gods of mythology!’ I tried to comprehend to what it was all pointing. My gaze dwelt, flickeringly, between the two. ‘If---’

An idea came swiftly, and I turned, and glanced rapidly upward, searching the gloomy crags, away to my left. Something loomed out under a great peak, a shape of greyness. I wondered I had not seen it earlier, and then remembered I had not yet viewed that portion. I saw it more plainly now. It was, as I have said, grey. It had a tremendous head; but no eyes. That part of its face was blank.

Now, I saw that there were other things up among the mountains. Further off, reclining on a lofty ledge, I made out a livid mass, irregular and ghoulish. It seemed without form, save for an unclean, half-animal face, that looked out, vilely, from somewhere about its middle. And then I saw others---there were hundreds of them. They seemed to grow out of the shadows. Several I recognized almost immediately as mythological deities; others were strange to me, utterly strange, beyond the power of a human mind to conceive.

On each side, I looked, and saw more, continually. The mountains were full of strange things---Beast-gods, and Horrors so atrocious and bestial that possibility and decency deny any further attempt to describe them. And I---I was filled with a terrible sense of overwhelming horror and fear and repugnance; yet, spite of these, I wondered exceedingly. Was there then, after all, something in the old heathen worship, something more than the mere deifying of men, animals, and elements? The thought gripped me---was there?

What were they, those Beast-gods, and the others? At first, they had appeared to me just sculptured Monsters placed indiscriminately among the inaccessible peaks and precipices of the surrounding mountains. Now, as I scrutinized them with greater intentness, my mind began to reach out to fresh conclusions. There was something about them, an indescribable sort of silent vitality that suggested, to my broadening consciousness, a state of life-in-death---a something that was by no means life, as we understand it; but rather an inhuman form of existence, that well might be likened to a deathless trance---a condition in which it was possible to imagine their continuing, eternally. ‘Immortal!’ the word rose in my thoughts unbidden; and, straightway, I grew to wondering whether this might be the immortality of the gods.

And then, in the midst of my wondering and musing, something happened. Until then, I had been staying just within the shadow of the exit of the great rift. Now, without volition on my part, I drifted out of the semi-darkness and began to move slowly across the arena---toward the House. At this, I gave up all thoughts of those prodigious Shapes above me---and could only stare, frightenedly, at the tremendous structure toward which I was being conveyed so remorselessly. Yet, though I searched earnestly, I could discover nothing that I had not already seen, and so became gradually calmer.

Presently, I had reached a point more than halfway between the House and the gorge. All around was spread the stark loneliness of the place, and the unbroken silence. Steadily, I neared the great building. Then, all at once, something caught my vision, something that came ’round one of the huge buttresses of the House, and so into full view. It was a gigantic thing, and moved with a curious lope, going almost upright, after the manner of a man. It was quite unclothed, and had a remarkable luminous appearance. Yet it was the face that attracted and frightened me the most. It was the face of a swine.

Silently, intently, I watched this horrible creature, and forgot my fear, momentarily, in my interest in its movements. It was making its way, cumbrously ’round the building, stopping as it came to each window to peer in and shake at the bars, with which---as in this house---they were protected; and whenever it came to a door, it would push at it, fingering the fastening stealthily. Evidently, it was searching for an ingress into the House.

I had come now to within less than a quarter of a mile of the great structure, and still I was compelled forward. Abruptly, the Thing turned and gazed hideously in my direction. It opened its mouth, and, for the first time, the stillness of that abominable place was broken, by a deep, booming note that sent an added thrill of apprehension through me. Then, immediately, I became aware that it was coming toward me, swiftly and silently. In an instant, it had covered half the distance that lay between. And still, I was borne helplessly to meet it. Only a hundred yards, and the brutish ferocity of the giant face numbed me with a feeling of unmitigated horror. I could have screamed, in the supremeness of my fear; and then, in the very moment of my extremity and despair, I became conscious that I was looking down upon the arena, from a rapidly increasing height. I was rising, rising. In an inconceivably short while, I had reached an altitude of many hundred feet. Beneath me, the spot that I had just left, was occupied by the foul swine-creature. It had gone down on all fours and was snuffing and rooting, like a veritable hog, at the surface of the arena. A moment and it rose to its feet, clutching upward, with an expression of desire upon its face such as I have never seen in this world.

Continually, I mounted higher. A few minutes, it seemed, and I had risen above the great mountains---floating, alone, afar in the redness. At a tremendous distance below, the arena showed, dimly; with the mighty House looking no larger than a tiny spot of green. The swine-thing was no longer visible.

Presently, I passed over the mountains, out above the huge breadth of the plain. Far away, on its surface, in the direction of the ring-shaped sun, there showed a confused blur. I looked toward it, indifferently. It reminded me, somewhat, of the first glimpse I had caught of the mountain-amphitheatre.

With a sense of weariness, I glanced upward at the immense ring of fire. What a strange thing it was! Then, as I stared, out from the dark center, there spurted a sudden flare of extraordinary vivid fire. Compared with the size of the black center, it was as naught; yet, in itself, stupendous. With awakened interest, I watched it carefully, noting its strange boiling and glowing. Then, in a moment, the whole thing grew dim and unreal, and so passed out of sight. Much amazed, I glanced down to the Plain from which I was still rising. Thus, I received a fresh surprise. The Plain---everything had vanished, and only a sea of red mist was spread far below me. Gradually as I stared this grew remote, and died away into a dim far mystery of red against an unfathomable night. A while, and even this had gone, and I was wrapped in an impalpable, lightless gloom.

\clearpage
